---
title: "Ranavirus simulations- Plotting BEAST outputs"
author: "Logan Billet"
date: "11/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
library(knitr)
library(lintr)
library(readxl)
```

Read in the data:
```{r}
data <- read_excel("BEAST_SimulationOutputs.xlsx")
data$sim.clockrate <- as.factor(data$sim.clockrate)
data$sim.rootheight.ybp <- as.factor(data$sim.rootheight.ybp)
levels(data$sim.rootheight.ybp) <- c("100 ybp", "1000 ybp")
data$sim.sequence.size <- as.factor(data$sim.sequence.size)
data$Num.taxa <- as.factor(data$Num.taxa)
data$sim.replicate <- as.factor(data$sim.replicate)
```


# Getting the data reformatted for plotting

I'm splitting things up into two plots- a 20 taxa plot and a 100 taxa plot. So, I'll make two separate data frames.

First, I'll need to subset the data by Num.taxa, and then by sequence size. This will allow me to plot the data with estimate as the y-axis, simulation clockrate as the x-axis, replicate as the grouping variable. Then I'll dodge the points by clockrate, and facet the data by root height.

First, we'll make the two 20 taxa datasets and plots.
```{r}
taxa_20 <- data %>%
  filter(Num.taxa == '20')

taxa_20_1200bp <- taxa_20 %>%
  filter(sim.sequence.size == "1200")

taxa_20_100000bp <- taxa_20 %>%
  filter(sim.sequence.size == "1e+05")
```



#BEAST simulation outputs - plotting the data


This is the plot for simulations with 20 taxa and 1200 bp nucleotide sequences
```{r}

taxa20_1200bp_plot <- ggplot(data=taxa_20_1200bp, aes(x= sim.clockrate, y= estimate.clockrate, group = sim.replicate, color=sim.clockrate)) +
  geom_point(position=position_dodge(width = 1))+
  labs(y = "Estimated substitution rate (log(subs/site/year))", x = "Simulated substitution rate (subs/site/year)", title = "20 taxa + 1200 bp nucleotide sequences")+
  facet_wrap(~sim.rootheight.ybp)+
  #theme(strip.background = element_blank(), strip.placement = "inside")+
  geom_errorbar(aes(ymin=clockrate.lower95, ymax=clockrate.upper95), position=position_dodge(width = 1))+
  scale_y_continuous(trans='log2', breaks=c(.00000001, .00000001, .0000001, .000001, .00001, .0001, .001))+
  geom_hline(yintercept=1e-04, linetype="dashed", color = "blue")+
  geom_hline(yintercept=1e-06, linetype="dashed", color = "green")+
  geom_hline(yintercept=1e-08, linetype="dashed", color = "red")+
  theme(axis.text.y = element_text(colour = "black"), axis.text.x=element_text(colour='black'))+
  theme_classic(base_size = 14) # theme for a nicer looking figure


taxa20_1200bp_plot +  theme(legend.title = element_blank())

```


This is the plot for simulations with 20 taxa and 100000 bp nucleotide sequences

```{r}

taxa20_100000bp_plot <- ggplot(data=taxa_20_100000bp, aes(x= sim.clockrate, y= estimate.clockrate, group = sim.replicate, color=sim.clockrate)) +
  geom_point(position=position_dodge(width = 1))+
  labs(y = "Estimated substitution rate (log(subs/site/year))", x = "Simulated substitution rate (subs/site/year)", title = "20 taxa + 100000 bp nucleotide sequences")+
  facet_wrap(~sim.rootheight.ybp)+
  #theme(strip.background = element_blank(), strip.placement = "inside")+
  geom_errorbar(aes(ymin=clockrate.lower95, ymax=clockrate.upper95), position=position_dodge(width = 1))+
  scale_y_continuous(trans='log2', breaks=c(.00000001, .00000001, .0000001, .000001, .00001, .0001, .001))+
  geom_hline(yintercept=1e-04, linetype="dashed", color = "blue")+
  geom_hline(yintercept=1e-06, linetype="dashed", color = "green")+
  geom_hline(yintercept=1e-08, linetype="dashed", color = "red")+
  theme(axis.text.y = element_text(colour = "black"), axis.text.x=element_text(colour='black'))+
  theme_classic(base_size = 14) # theme for a nicer looking figure


taxa20_100000bp_plot +  theme(legend.title = element_blank())
```


Then, we'll make the 100 taxa datasets and plots
```{r}
taxa_100 <- data %>%
  filter(Num.taxa == '100')

taxa_100_1200bp <- taxa_100 %>%
  filter(sim.sequence.size == "1200")

taxa_100_100000bp <- taxa_100 %>%
  filter(sim.sequence.size == "1e+05")
```
This is the plot for simulations with 100 taxa and 1200 bp nucleotide sequences
```{r}

taxa100_1200bp_plot <- ggplot(data=taxa_100_1200bp, aes(x= sim.clockrate, y= estimate.clockrate, group = sim.replicate, color=sim.clockrate)) +
  geom_point(position=position_dodge(width = 1))+
  labs(y = "Estimated substitution rate (log(subs/site/year))", x = "Simulated substitution rate (subs/site/year)", title = "100 taxa + 1200 bp nucleotide sequences")+
  facet_wrap(~sim.rootheight.ybp)+
  #theme(strip.background = element_blank(), strip.placement = "inside")+
  geom_errorbar(aes(ymin=clockrate.lower95, ymax=clockrate.upper95), position=position_dodge(width = 1))+
  scale_y_continuous(trans='log2', breaks=c(.00000001, .00000001, .0000001, .000001, .00001, .0001, .001))+
  geom_hline(yintercept=1e-04, linetype="dashed", color = "blue")+
  geom_hline(yintercept=1e-06, linetype="dashed", color = "green")+
  geom_hline(yintercept=1e-08, linetype="dashed", color = "red")+
  theme(axis.text.y = element_text(colour = "black"), axis.text.x=element_text(colour='black'))+
  theme_classic(base_size = 14) # theme for a nicer looking figure


taxa100_1200bp_plot +  theme(legend.title = element_blank())

```


This is the plot for simulations with 100 taxa and 100000 bp nucleotide sequences
```{r}

taxa100_100000bp_plot <- ggplot(data=taxa_100_100000bp, aes(x= sim.clockrate, y= estimate.clockrate, group = sim.replicate, color=sim.clockrate)) +
  geom_point(position=position_dodge(width = 1))+
  labs(y = "Estimated substitution rate (log(subs/site/year))", x = "Simulated substitution rate (subs/site/year)", title = "100 taxa + 100000 bp nucleotide sequences")+
  facet_wrap(~sim.rootheight.ybp)+
  #theme(strip.background = element_blank(), strip.placement = "inside")+
  geom_errorbar(aes(ymin=clockrate.lower95, ymax=clockrate.upper95), position=position_dodge(width = 1))+
  scale_y_continuous(trans='log2', breaks=c(.00000001, .00000001, .0000001, .000001, .00001, .0001, .001))+
  geom_hline(yintercept=1e-04, linetype="dashed", color = "blue")+
  geom_hline(yintercept=1e-06, linetype="dashed", color = "green")+
  geom_hline(yintercept=1e-08, linetype="dashed", color = "red")+
  theme(axis.text.y = element_text(colour = "black"), axis.text.x=element_text(colour='black'))+
  theme_classic(base_size = 14) # theme for a nicer looking figure


taxa100_100000bp_plot +  theme(legend.title = element_blank())

```















